Optimal analysis of Discrete-time Time-Invariant Affine Systems

### 2018-09-27 paper.tex
new

\section{Introduction}
Some catastrophic events~\cite{johnson2005natural,mcquaid2012software} caused by bugs in programs explain the importance of the formal verification of programs~\cite{cousot2010gentle}.
Furthermore, since the success of Astrée~\cite{delmas2007astree,souyris2007experimental,bouissou2009space}, industrials are  more and more concerned with formal verification.  
%Goal of verification : some catastrophic events caused by bugs in programs leaded to a new branch in computer science called formal verification of programs\ForAuthors{trouver une citation sur les premiers travaux en verification}.

Formal verification provides a deeper analysis compared to traditional test techniques. Indeed, tests cannot cover all possible situations. The principle of formal verification consists in an exhaustive approach to prove properties on programs. Various techniques exist but none of them cannot be complete. This means that they cannot give a validation of the property for every program. 
In this paper, we are interested in static analysis approach. It consists in verifying the system without using simulations and prove the property from the dynamics initial conditions.

Some works propose numerical optimization techniques to solve verification problem.

Numerical programs in an ideal world can be viewed as a discrete-time dynamical systems. Some numerical
programs comes from discretization or linearization techniques. Thus linear systems forms a huge class of interesting problems.

During the last twenty years,  the verification of linear systems techniques evolve and propose new challenges. Classical approaches come from theoretical computer science. Those approaches are based on static analysis such as model-checking or abstract interpretation. 
The goal of the paper is to offer a complete analysis of discrete-time linear systems.
Indeed, the static analysis of such systems based on abstract interpretation can prove 
valid properties. However, because abstractions and the presence of false alarms, the challenge which consists in invalidating properties remains extremely difficult for analyzers. The problem addressed here can be reduced to maximize a function over the reachable values set of a dynamical system.
The method developed here consists in replacing an infinite numbers of maximization problems by a finite number of them.  We propose to compute an integer after which we know that the optimal value cannot be reached.

@@

\paragraph{Motivations}
Motivations come from computer science and static analysis. More precisely from abstract interpretation 
\paragraph{Related works}
Verification linear /affine systems
Optimization tools for verification
Formal methods tools 
Control problems with reachability
\paragraph{Contributions}
Compute an integer that permits to solve a optimization problem with reachability constraints.
This problem is then solved exactly
From the computer science point of view, the motivation, we can prove or disprove a property
\paragraph{Organization of the paper}
Problem statement, the main results, the computation of the integer, the experimentation and finally a conclusion.
\paragraph{Notations}
In this paper, we denote by $\Idd$ the identity matrix of size $d\times d$. We denote, for a matrix $M$, by $M^\intercal$ the transpose of $M$. The set of the (square) symmetric matrices  (the matrices $M$ such that $M=M^\intercal$) of size $d$ is denoted by $\sym$.

>>>>

\section{Introduction}
\paragraph{Motivations.}
Some catastrophic events~\cite{johnson2005natural,mcquaid2012software} caused by bugs in programs explain the importance of the formal verification of programs in general and static analysis of programs~\cite{cousot2010gentle} in particular. The interest is shared by both theorists and industrials. Indeed, since the success of Astrée~\cite{delmas2007astree,souyris2007experimental,bouissou2009space}, industrials are  more and more concerned with static analysis.  To solve static analysis problems is thus an interesting scientific challenge for the theory as well as for the applications and the pratical  side.

This interest for static analysis techniques comes from its exhaustivity. Indeed static analysis offers a deeper analysis compared to traditional test techniques which cannot cover all possible situations. The principle of static analysis consists in proving properties automatically on programs without executing them. Static analyzers only use the structures of programs not the values generated by the program.  Numerous various techniques exist but none of them cannot be complete. This means that no program can validate or disprove a given property for all programs. Static analysis is not only useful for programs but can also be applied to validate properties on dynamical systems in general. The principle is still the same : to prove properties on dynamical systems without any simulations.

To prove or disprove property even for discrete-time linear system is still an open challenge~\cite{acceleration-popl14,7403149,DBLP:conf/icalp/AlmagorCO018}.  The difficulty of the analysis lies in the infinite number of points in the reachable values set. Static analysis abstract interpretation~\cite{cousot2001abstract} based proposes to prove properties by means of over-approximations of reachable values set. Then abstract interpretation has difficulties to disprove a property since over-approximations can contain non-reachable values which violate the property. 
%Static analysis is mainly divided into two approaches the model-checking and abstract interpretation. Model-checking exploits finite models to disprove property by the construction of a counterexemple whereas abstract interpretation proposes to prove a property by means of over-approximation of reachable values set. Those techniques are then limited since model-checking has difficulties to prove properties and abstract interpretation has difficulties to disprove a property.

The main motivation of the paper is to propose a technique that can be, for a given discrete-time affine system and a given property, to prove or disprove the property on the system.

%To do this, we reformulate the verification problem as an optimization problem with dynamical system contraints. This optimization problem is then solved exactly. The exact optimal solution can guarantee whether the property holds or not. Note that this particular subclass of optimization problem is called robust-to-dynamics optimization in~\cite{7403149}.   
\paragraph{Contributions}
Compute an integer that permits to solve a optimization problem with reachability constraints.
This problem is then solved exactly
From the computer science point of view, the motivation, we can prove or disprove a property

\paragraph{Related works .}
For the static analysis of discrete-time affine systems, numerous works in computer science or control theory exist. 
in computer science as we said earlier, works are decomposed into two world model checking community and abstract interpretation.

Model-checking approach as said earlier can disprove a property by the synthesis of counter-examples. The direct validation of the property would require an infinite number of runs.

Proof techniques based on Lyapunov functions~\cite{roozbehani2013optimization,blanchini2009lyapunov} automatically deals with non-convexities. This makes difficult the exact resolution of the verification problem. To use Lyapunov functions to solve verification problems lead  to over-approximations and only conclusion that can be made is "the property holds or we do not know".

In~\cite{adje2015property}, the authors propose  an approach based on Sums-of-Squares optimization to prove property on discrete-time polynomial systems. The verification problem is reduced to an optimization problem exactly the same that we use here.  The objective function is provided from the property to prove. The proof of the property pass through the synthesis of a polynomial invariant i.e.  a sublevel set of a polynomial that over-approximates the reachable value set of the polynomial system. The computed sublevel is the one that minimizes the over-approximation. This approach uses abstract interpretation and then cannot disprove a property. Actually, the main issue is that we cannot prove that there is a "zero duality gap result" . The same kind of approach exists for piecewise affine systems~\cite{10.1007/978-3-319-54292-8_2} with the same disadvantage.

The closest work seems to be the work of Ahmadi et al~\cite{7403149,ahmadi2018robust} which introduces the class of robust-to-dynamics optimization problem. Ahmadi et al solve the problem where the dynamics and the objective function are linear. Moreover, they need a bounded invariant set to solve their problem. In our work, we deal with non-linear quadratic  forms and affine dynamical systems (to pass from linear to affine seems to be a very small improvment).  Moreover, we do not need a bounded invariant to solve our robust-to-dynamics optimization problem.  Nevertheless, some similarities exist between their approach and the one proposed here : the use of a solution of the discrete Lyapunov matrix equation and the computation of a the number of steps to convergence.

%In this paper, we are interested in proving properties on discrete-time affine systems. The proof is based on a static analysis approach. It consists in verifying the dynamical system without using simulations and prove the property from the dynamics and the initial conditions.  We are focused on numerical properties. Properties considered can be the boundedness of some state-variables, to belong to some safe set or to avoid an unsafe set.

%Some works propose numerical optimization techniques to solve verification problem.
%Numerical programs in an ideal world can be viewed as a discrete-time dynamical systems. Some numerical
%programs comes from discretization or linearization techniques. Thus linear systems forms a huge class of interesting problems.

%During the last twenty years,  the verification of linear systems techniques evolve and propose new challenges. Classical approaches come from theoretical computer science. Those approaches are based on static analysis such as model-checking or abstract interpretation. 
%The goal of the paper is to offer a complete analysis of discrete-time linear systems.
%Indeed, the static analysis of such systems based on abstract interpretation can prove 
%valid properties. However, because abstractions and the presence of false alarms, the challenge which consists in invalidating properties remains extremely difficult for analyzers. The problem addressed here can be reduced to maximize a function over the reachable values set of a dynamical system.
%The method developed here consists in replacing an infinite numbers of maximization problems by a finite number of them.  We propose to compute an integer after which we know that the optimal value cannot be reached.

@@

Verification linear /affine systems
Optimization tools for verification
Formal methods tools 
Control problems with reachability

\paragraph{Organization of the paper}
Problem statement, the main results, the computation of the integer, the experimentation and finally a conclusion.
\paragraph{Notations}
In this paper, we denote by $\Idd$ the identity matrix of size $d\times d$. We denote, for a matrix $M$, by $M^\intercal$ the transpose of $M$. The set of the (square) symmetric matrices  (the matrices $M$ such that $M=M^\intercal$) of size $d$ is denoted by $\sym$.
